### 2025-05-24 text/50-ExperimentalMethod.qmd

## Experimental method {#methodology}
We can now detail our experimental method. We evaluate the standard parametric approach (*PAR*) and the three rank-transformation methods (*RNK*, *INT*, and *ART*) that we introduced earlier. We conduct a series of Monte Carlo experiments that assess their performance under a variety of experimental configurations:

1. We evaluate *ratio* and *ordinal* data. For ratio data, we examine four representative *continuous* distributions (normal, log-normal, exponential, and Cauchy distribution) and two *discrete* distributions (Poisson and binomial distribution). For ordinal data, we examine distributions for 5-level, 7-level, and 11-level Likert item responses.

2. We present results for five experimental designs. To simplify our presentation, we start with a 4 $\times$ 3 repeated-measures factorial design. We then show how our conclusions generalize to four additional designs: (i) a 2 $\times$ 3 between-subjects design; (ii) a 2 $\times$ 4 mixed design, with a between-subjects factor and a repeated-measures factor; (iii) a 2 $\times$ 2 $\times$ 2 repeated-measures design, and (iv) a 3 $\times$ 3 $\times$ 3 repeated-measures design.

3. We evaluate three sample sizes, $n=10$, $n=20$, and $n=30$, where $n$ represents the cell size in an experimental design. For within-subjects designs where all factors are treated as repeated measures, $n$ coincides with the number $N$ of subjects (or commonly human participants in HCI research). In contrast, for a 2 $\times$ 3 between-subjects design, if the cell size is $n = 20$, then the number of subjects is $N = 120$.

4. We test the robustness of the methods under unequal variances.

5. In addition to Type I error rates, we compare the statistical power of the methods and also evaluate the quality of their effect size estimates.

6. We assess the above measures for both main and interaction effects and examine how growing effects on one or two factors affect the Type I error rate on other factors and their interactions.

<!--
Furthermore, we assess how a growing interaction effect, either in isolation or in combination with a main effect, can affect the inference of main effects. -->

Previous evaluations of rank transformation methods [@Beasley:2009; @luepsen:2018] have also tested unbalanced designs, where the cell size is not constant across all levels of a factor. When combined with unequal variances, unbalanced designs are often problematic for both parametric procedures [@blanca:2018] and rank transformation methods [@Beasley:2009; @luepsen:2018]. As we mentioned earlier, we do not evaluate unbalanced designs here. However, we provide additional experimental results on missing data in the [appendix](appendix.html).

### Statistical modeling
To model the latent variable $Y$, we use a two-way (two factors) or a three-way (three factors) mixed-effects model. For simplicity, we explain here the model for two factors. Its extension to three factors is straightforward. The model has the following form:

$$ 
y_{ijk} = \mu + s_k + a_1 x_{1i} + a_2 x_{2j} + a_{12} x_{1i} x_{2j} + \epsilon_{ijk}
$${#eq-linear-model}

 - $\mu$ is the grand mean

 - $s_k$ is the random intercept effect of the *k*-th subject, where $k = 1..n$

 - $x_{1i}$ is a numerical encoding of the *i*-th level of factor $X_1$, where $i = 1..m_1$

 - $x_{2j}$ is a numerical encoding of the *j*-th level of factor $X_2$, where $j = 1..m_2$

 - $a_1$, $a_2$, and $a_{12}$ express the magnitude of main and interaction effects

 - $\epsilon_{ijk}$ is the experimental error effect

To encode the levels of the two factors $x_{1i} \in X_1$ and $x_{2j} \in X_2$ we proceed as follows:

1. We normalize the distance between their first and their second levels such that $x_{12} - x_{11} = 1$ and $x_{22} - x_{21} = 1$. This approach enables us to conveniently control for the main and interaction effects by simply varying the parameters $a_1$, $a_2$, and $a_{12}$.

2. For the remaining levels, we randomly sample from a uniform distribution that spans the range between these two extreme levels, i.e., between $x_{11}$ and $x_{12}$ for $X_1$, and between $x_{21}$ and $x_{22}$ for $X_2$. This approach allows us to generate and evaluate a substantial variety of configurations, each representing different relative effects between levels.

3. We require all levels to sum up to 0, or $\sum\limits_{i=1}^{m_1}  x_{1i} = 0$ and $\sum\limits_{j=1}^{m_2}  x_{2j} = 0$, which ensures that the grand mean is $\mu$.

For example, we can encode a factor with four levels as $\{-.6, .4, .1, .1\}$ or as $\{-.5, .5, .3, -.3\}$.

While random slope effects can have an impact on real experimental data [@barr:2013], we do not consider them here for two main reasons: (1) to be consistent with previous evaluations of the ART procedure [@elkin:2021]; and (2) because mixed-effects procedures with random slope effects are computationally demanding, adding strain to simulation resources. However, there is no good reason to believe that adding random slope effects would impact our findings and conclusions.

### Population control and distribution conversions 
To simplify our simulations, we fix the following population parameters: $\mu = 0$, $\sigma = 1$, and $\sigma_s = 1$. We then control the magnitude of effects by varying $a_1$, $a_2$, and, for some experiments $a_{12}$. @fig-effects presents the range of values that we test for $a_1$ and $a_2$. We also visualize their effects on the latent variable for factors with two, three, and four categorical levels.

::: {#fig-effects}
```{r, echo=FALSE, message=FALSE, fig.height=3, fig.width = 9, warning=FALSE}
source("effects_plot.R")
plotEffects()
```
Varying $a_1$ or $a_2$ to control the magnitude of main effects. The plots show examples of population distributions on the latent variable $Y$ for factors ($X_1$ or $X_2$) with two (top), three (middle), or four categorical levels (bottom).    
:::

We follow the approach of @DeBruine:2021 and use the R package *faux* v1.2.1 [@faux] to simulate data for our mixed-effects models. We assume that the distributions of random intercepts and errors are normal, or $s_k \sim N(0,\sigma_s)$ and $\epsilon_{ijk} \sim N(0,\sigma)$.
To simulate the observed variable $Y'$, we then transform the response values $y_{ijk}$ as described in [Section 3](#approach). A key advantage of this method is that we can generate responses for any distribution, while we control effects on the latent variable in the exact same way.

### Implementation of rank transformation methods 
For the aligned rank transformation (ART), we use the R implementation of ARTool v0.11.1 [@artool]. For the pure rank transformation (RNK), we use R's *rank()* function. We use the *Rankit* formulation [@Bliss:1956] for the inverse normal transformation (INT), as explained earlier. Unless explicitly mentioned otherwise, we use the formula ``Y' ~ X1*X2 + (1|S)`` with R's *lmer* function, except for the between-subjects design where we use the formula ``Y' ~ X1*X2 + Error(S)`` with R's *aov* function.

### Evaluation measures
Significance tests have two types of errors. *Type I errors*, or false positives, are mistaken rejections of the null hypothesis. Type II errors, of false negatives, are failures to reject a null hypothesis that is actually true. In our illustrative example in @fig-example, a Type I error is finding that there is an effect of the choice of the technique on time performance. A *Type II error*, in turn, is finding that the task difficulty has no effect on time performance.

Statistical significance testing requires setting a significance threshold known as significance or $\alpha$ (alpha) level, with typical values $\alpha = .05$ and $\alpha = .01$. The Type I error rate of a well-behaved significance test should be close to this nominal alpha level. An error rate clearly above this level suggests that the significance test is too liberal, while an error rate clearly below this level suggests that the test is too conservative. Four of our experiments specifically assess the Type I error rate of the methods. We test two significance levels: $\alpha = .05$ and $\alpha = .01$. For brevity, we only report results for $\alpha = .05$ in the main paper and include additional results in our supplementary material.

We do not directly evaluate Type II errors. Instead, we report on statistical *power* defined as $Power = 1 - \beta$, where $\beta$ is the rate of Type II errors. Significance tests do not provide any power guarantees. However, we can compare the power of different methods to evaluate their relative performance. In addition to power, we assess effect size estimates, where we use as ground truth the estimates of a parametric ANOVA conducted on the latent variable $Y$. While partial $\eta^2$ is the most commonly used effect size measure, we chose to base our analysis on Cohen's $f$, as its expected value is proportional to the real magnitude of effect. However, $\eta^2$ can be directly derived from Cohen's $f$ as follows:

$$ 
\eta^2 = \frac{f^2}{1 + f^2}
$${#eq-cohensf}

As we explained earlier, interaction effects are deformed when we transform the latent variable $Y$ to obtain the observed responses. Such transformations also affect the Type I error rates and power that we observe. Here, we are interested in evaluating these measures with respect to the effects we apply to the latent variable $Y$. We make this distinction whenever it is relevant in our analysis.

### Hardware platform and iterations
Our experimental R code is available in our supplementary material. We ran our experiments separately in a cluster of 8 machines Dell R640 Intel Xeon Silver 4112 2.6GHz with 4 cores and 64 GB memory. Our R code was parallelized to use all four cores of each machine. Some experiments took a few hours to complete, while others took several days.

To estimate the power and Type I error rates of the four methods with enough precision, we ran $5000$ iterations for each population configuration and each sample size.

<!--
We evaluate the Type I error rate, that is the rate of false positives of each method. To evaluate the Type I error rate for the interaction effect, we set $a_{12} = 0$. Likewise, to evaluate the Type I error rate for the second factor $X_2$, we set $a_{2} = 0$. For each population configuration and sample size $n$, we run $5000$ iterations and test two significance levels: $\alpha = .05$ and $\alpha = .01$. For brevity, we only report results for $\alpha = .05$ in the main paper, while additional results are presented in supplementary materialsxa
-->