### 2025-12-12 appendix_I.qmd
working on appendix I

---
title: "The illusory promise of the Aligned Rank Transform"
subtitle: "Appendix I. Mathematical explanations and proofs"
author: 
  - name: Theophanis Tsandilas
    orcid: 0000-0002-0158-228X
    email: theophanis.tsandilas@inria.fr
    affiliations:
      - name: Université Paris-Saclay, CNRS, Inria, LISN
        country: France
  - name: Géry Casiez
    orcid: 0000-0003-1905-815X
    email: gery.casiezuniv-lille.fr
    affiliations:
      - name: Univ. Lille, CNRS, Inria, Centrale Lille, UMR 9189 CRIStAL
        country: France
bibliography: bibliography.bib
csl: canadian-journal-of-philosophy.csl

tbl-cap-location: top
---

## Comparing Type I error rates under non-linear models
Consider again the general equation that connects the expected value of a response variable $Y$, which may follow a non-normal distribution, to the predictors in an experimental design with two factors $x_1$ and $x_2$:

$$
E(Y|x_1, x_2) = g^{-1}(a_0 + a_1 x_1 + a_2 x_2 + a_{12}x_1 x_2)
$$
where $g^{-1}$ is the inverse link function.

We explain why in the main scenarios in this article, the interpretation of effects is not affected by the way the null hypothesis is defined. Our analysis is largely based on related analyses by @Ai:2003, @Greene:2010, and @McCabe:2022.

To simplify our analysis, we use $\eta(x_1, x_2)$ to denote the linear part of the model:

$$
\eta(x_1, x_2) = a_0 + a_1 x_1 + a_2 x_2 + a_{12}x_1 x_2
$$

We also use $\mu(x_1, x_2)$ to denote the expected value as a function of $x_1$ and $x_2$:

$$
\mu(x_1, x_2) = E(Y|x_1, x_2) = g^{-1}(\eta(x_1, x_2))
$$

### Main effects
We first show that there is no ambiguity in the interpretation of main effects when $a_{12}=0$.

Without loss of generality, we assume that $a_1$ is non-zero and focus on the observed effect on the second factor $x_2$ when $a_2 = 0$. Then: 
$$
\begin{aligned}
\eta(x_1, x_2) = a_0 + a_1 x_1, \qquad
\mu(x_1, x_2) = g^{-1}(a_0 + a_1 x_1)
\end{aligned}
$$

We consider two different situations:

(a) *Continuous $x_2$ --- partial-derivative definition of the effect*. In this case, we can define the effect of $x_2$ on the response as the partial derivative $\frac{\partial \mu}{\partial x_2}$. This becomes:
$$\frac{\partial \mu(x_1, x_2)}{\partial x_2}  = \frac{\partial}{\partial x_2} g^{-1}(a_0 + a_1 x_1) = 0$$
since the term $a_0 + a_1 x_1$ does not contain $x_2$. Thus, $x_2$ has no effect on the responses.

(b) *Categorical $x_2$ --- contrast definition of the effect*. This case is closer to the scenarios we consider in our simulations, where all independent variables are categorical. Let $L_2$ be the categorical levels of $x_2$. Then, the effect of $x_2$ can be defined as the difference in means across its levels for a fixed $x_1$:
$$\Delta_{u,\upsilon}E(Y|x_1, x_2) = \mu(x_1, u) - \mu(x_1, \upsilon) = g^{-1}(a_0 + a_1 x_1) - g^{-1}(a_0 + a_1 x_1) = 0$$
for any two levels $u, \upsilon \in L_2$.

**Conclusion:** Let $a_{12} = 0$. If $a_2 = 0$ then the effect on the mean of responses is also zero. Thus, whether we define Type I errors in terms of the coefficient $a_2$ or in terms of the means of the responses, results should not change. We emphasize that this result **is not specific to means**. If $Y$ depends on $x_1$, $x_2$ only through the predictor $\eta(x_1, x_2)$ --- that is, variances or other distribution shape parameters do not depend on these variables --- then no effect of $x_2$ can appear on the mean, the median, or any other quantile of the distribution when both $a_{12}$ and $a_2$ are zero.

**Special note:** However, if $a_{12}$ is non-zero, it is possible to observe an effect of $x_2$ on responses even if $a_2 = 0$.

### Interactions
We similarly show that there is no ambiguity in the interpretation of the interaction effect when either $a_1 = 0$ or $a_2 =0$. Without loss of generality, we assume below that $a_2 = 0$ while $a_1$ is non-zero. For $a_{12}=0$: 
$$
\begin{aligned}
\eta(x_1, x_2) = a_0 + a_1 x_1, \qquad
\mu(x_1, x_2) = g^{-1}(a_0 + a_1 x_1)
\end{aligned}
$$

A common definition of interaction on the response scale [@Ai:2003] is that the *marginal effect* of $x_1$ depends on $x_2$. Again, we consider two situations:

(a) *Continuous $x_1$ and $x_2$ --- definition of interaction through second-order cross-partial derivative*. For continuous $x_1$, this is expressed by the second-order cross-partial derivative $\frac{\partial^2 \mu}{\partial x_2 \partial x_1}$, which becomes:
 $$\frac{\partial^2 \mu(x_1, x_2)}{\partial x_2 \partial x_1}  = \frac{\partial}{\partial x_2} [\dot{g}^{-1}(\eta(x_1, x_2)) \cdot a_1] = \ddot{g}^{-1}(\eta(x_1, x_2)) \cdot \frac{\partial \eta(x_1, x_2)}{\partial x_2} \cdot a_1$$ 
 based on the rule of chain, where $\dot{g}^{-1}$ and $\ddot{g}^{-1}$ are the first and second derivatives of the inverse link function (assuming that $g^{-1}$ is twice differentiable). But we have:
 $$\frac{\partial \eta(x_1, x_2)}{\partial x_2} = a_2 + a_{12}x_1 = 0$$ because $a_2$ and $a_{12}$ are both zero. Thus, $\frac{\partial^2 \mu}{\partial x_2 \partial x_1}$ becomes zero, which means that the marginal effect (slope) of $x_1$ does not change with $x_2$ (and vice versa). There is therefore no interaction on the response scale.

(b) *Categorical $x_1$ and $x_2$ --- definition of interaction through double discrete difference*. Let $L_1$ and $L_2$ be the categorical levels of $x_1$ and $x_2$, respectively. Then, the interaction effect can be defined as the double difference in means across all combinations of levels in $L_1$ and $L_2$:
$$\Delta_{u_1, \upsilon_1} \Delta_{u_2, \upsilon_2}E(Y|x_1, x_2) = [\mu(u_1,u_2) - \mu(u_1, \upsilon_2)] - [\mu(\upsilon_1,u_2) - \mu(\upsilon_1, \upsilon_2)] $$
for any levels $u_1, \upsilon_1 \in L_1$ and $u_2, \upsilon_2 \in L_2$.

**Conclusion:** Let $a_{12} = 0$. If $a_2 = 0$ then the effect on the mean of responses is also zero. Thus, whether we define Type I errors in terms of the coefficient $a_2$ or in terms of the means of the responses, results should not change.

## References {.unnumbered}

::: {#refs}
:::