### 2025-12-18 appendix_II.qmd

---
title: "The illusory promise of the Aligned Rank Transform"
subtitle: "Appendix II. Analytical derivations and mathematical proofs"
author: 
  - name: Theophanis Tsandilas
    orcid: 0000-0002-0158-228X
    email: theophanis.tsandilas@inria.fr
    affiliations:
      - name: Université Paris-Saclay, CNRS, Inria, LISN
        country: France
  - name: Géry Casiez
    orcid: 0000-0003-1905-815X
    email: gery.casiezuniv-lille.fr
    affiliations:
      - name: Univ. Lille, CNRS, Inria, Centrale Lille, UMR 9189 CRIStAL
        country: France
bibliography: bibliography.bib
csl: canadian-journal-of-philosophy.csl

tbl-cap-location: top
---

## Absence of spurious effects when $a_2=0$ and $a_{12} = 0${#effects-proof}
Let $x_1$ and $x_2$ be two independent variables, and suppose the conditional distribution of the response $Y | (x_1, x_2)$ depends on these variables only through the linear predictor:
$$
\eta(x_1, x_2) = a_0 + a_1 x_1 + a_2 x_2 + a_{12}x_1 x_2
$$

Assume that the conditional mean can be written as:
$$
\mu(x_1, x_2) = E(Y|x_1, x_2) = h(\eta(x_1, x_2))
$$
where $h$ is a monotone function. This function may coincide with the inverse link function $g^{-1}$ in a GLM, but it may also arise from other frameworks, including transformation models or latent-variable constructions. Thus it does not necessarily need to be smooth or invertible; for example, this assumption is not required for the ordinal example discussed later.

Suppose $a_2 = 0$ and $a_{12} = 0$. Then:

1. The independent variable $x_2$ has no effect on the response scale.
2. There is no interaction between $x_1$ and $x_2$ on the response scale.

Thus, whether Type I errors are defined in terms of the model coefficients $a_2$ (when $a_{12} = 0$) or $a_{12}$ (when $a_2 = 0$) or in terms of the patterns of observed means (and alternatively medians or other quantiles) on the response scale makes no difference.

### Proof
Under $a_2 = a_{12} = 0$, the linear predictor reduces to

$$\eta(x_1, x_2) = a_0 + a_1x_1$$

which does not depend on $x_2$. Hence, the conditional distribution $Y | (x_1, x_2)$, including its mean, median, and all quantiles, is invariant to $x_2$. This establishes the absence of a main effect of $x_2$ or an effect of its interaction $x_2$.

We further explain this results for special cases, where we largely rely on the relevant discussions of @Ai:2003, @Greene:2010, and @McCabe:2022.

#### Main effect of $x_2$
If $x_2$ is continuous and $h$ is differentiable, the effect of $x_2$ on the response can be defined as:

$$\frac{\partial \mu(x_1, x_2)}{\partial x_2}  = h'(\eta) \frac{\partial \eta}{\partial x_2} = 0$$
where $h'$ denotes the first derivative of $h$.

If $x_2$ is categorical, all contrasts across its levels are zero because the mean response is identical at each level:
$$\Delta_{u,\nu}\mu(x_1, x_2) = \mu(x_1, u) - \mu(x_1, \nu) = h(a_0 + a_1 x_1) - h(a_0 + a_1 x_1) = 0$$
where $u, \nu$ are any two levels of $x_2$.

#### Interaction between $x_1$ and $x_2$
An interaction on the response scale is present if the marginal effect of $x_1$ depends on $x_2$. For continuous predictors and assuming that $h$ is twice differentiable, this is characterized by the mixed partial derivative $\frac{\partial^2 \mu}{\partial x_2 \partial x_1}$.

Using the chain rule:

$$\frac{\partial^2 \mu(x_1, x_2)}{\partial x_2 \partial x_1} = h''(\eta) \frac{\partial \eta}{\partial x_2} \frac{\partial \eta}{\partial x_1} + h'(\eta)\frac{\partial^2\eta}{\partial x_2 \partial x_1} = 0$$

Thus the marginal effect of $x_1$ does not vary with $x_2$, and there is no interaction.

For categorical predictors, the analogue of a mixed partial derivative is the double difference:

$$\Delta_{u_1, \nu_1} \Delta_{u_2, \nu_2}\mu(x_1, x_2) = [\mu(u_1,u_2) - \mu(u_1, \nu_2)] - [\mu(\nu_1,u_2) - \mu(\nu_1, \nu_2)]=$$ 
$$= [h(a_0 + a_1 u_1) - h(a_0 + a_1 u_1)] - [h(a_0 + a_1 \nu_2) - h(a_0 + a_1 \nu_1)] = 0$$

for any levels $u_1, \nu_1$ of $x_1$ and $u_2, \nu_2$ of $x_2$. Again, there is no observed interaction.

### Illustrative examples
We discuss two examples.

#### Lognormal data
Let $x_1$ and $x_2$ be two experimental factors. Suppose the true model of the data is
$$
\begin{aligned}
log Y = a_0 + a_1 x_1 + \varepsilon, \qquad
\varepsilon \sim \mathcal{N}(0, \sigma^2)
\end{aligned} 
$$
so that 
$$Y | (x_1, x_2) \sim Lognormal(a_0 + a_1 x_1, \sigma^2)$$

Since the conditional distribution of $Y$ does not depend on $x_2$, there is no main effect of $x_2$ on the mean, median, or any quantile of the response distribution. In particular, the conditional mean
$$\mu(x_1,x_2) = E(Y|x_1, x_2) = exp(a_0 + a_1 x_1 + \frac{1}{2}\sigma^2)$$
remains identical across all levels of $x_2$.

For the same reasons, there is no interaction on the response scale.

#### Ordinal responses
Suppose there exists a latent continuous variable:

$$Y^* = a_0 + a_1 x_1 + \varepsilon$$

and the observed response $Y$ is obtained by applying fixed thresholds: 
$$
\begin{aligned}
Y = k\ \ \text{if}\ \ \tau_{k-1} < Y^* \leq \tau_k, 
\qquad k = 1,2 ... K
\end{aligned} 
$$ 
Again, the conditional distribution $Y| (x_1, x_2)$ depends on the independent variables only through $\eta(x_1, x_2)=a_0+a_1 x_1$, and there is no effect of $x_2$ on either the latent or the observed ordinal scale.

In this case, the mapping from $\eta$ to the observed response is monotone but not strictly monotone and is therefore non-invertible. As a result, a true effect of $x_2$ on the latent variable (i.e., $a_2 \neq 0$) may be attenuated or entirely unobservable on the ordinal response scale. This does not affect Type I error rates, provided the null hypothesis is defined on the latent scale and the thresholds are fixed, because the absence of a latent effect implies the absence of an effect on the observed scale. However, power and effect sizes can be reduced when effects are assessed on the ordinal scale, reflecting the loss of information due to discretization.

### Invariance under monotone transformations 
Suppose we apply a monotone transformation $T$ directly to the observed responses---such as the rank transformation (RNK) or the inverse normal transformation (INT). Because $T$ is monotone, it preserves the ordering of observations within each condition. The transformed conditional distribution $T(Y) | (x_1, x_2)$ therefore depends on the independent variables only through $\eta(x_1, x_2) = a_0+a_1 x_1$. Consequently, the mean, median, and all quantiles of $T(Y) | (x_1, x_2)$ remain invariant to $x_2$, and marginal effects of $x_1$ do not depend on $x_2$.

We emphasize that ART is not a monotone transformation. Instead of transforming $Y$ directly, ART first constructs aligned responses by subtracting and adding sample means on the response scale. These alignment terms can vary across levels of $x_2$ or combinations of $(x_1, x_2)$. Therefore, ART is not invariant to nonlinear mean–response relationships.

<!--
In conclusion, $T$ does not introduce a main effect of $x_2$ or an interaction between $x_1$ and $x_2$ when $a_2=0$ and $a_{12}=0$.

### ART introducing spurious effects 
ART differs fundamentally from simple monotone transformations. Instead of transforming $Y$ directly, ART first constructs aligned responses by **subtracting and adding sample means on the response scale**, for example:

$$Y^{(B)} = Y - \bar{Y}_{ij} + \bar{Y}_{.j} - \bar{Y}_{..}$$
or analogous expressions for interactions, and only then applies ranking. The conditional distribution $Y^{(B)} | (x_1, x_2)$ now depends on $x_2$ through the arithmetic mean components even though $a_2 = 0$ and $a_{12}=0$.

These alignment terms can vary across levels of $x_2$ due to sampling variability alone, and depending on $h$, they can acquire systematic shifts across levels of $x_2$ or combinations of $(x_1, x_2)$, even when the true conditional distribution does not. in particular, the term $\bar{Y}_{.j} - \bar{Y}_{..}$ is constant within each level of $x_2$ and shifts all aligned observations for that level either upward or downward.

After ranking, observations from one level of $x_2$ tend to receive systematically higher or lower ranks. Our results confirm this problem.
-->

## References {.unnumbered}

::: {#refs}
:::