### 2025-12-20 text/62-Results-AcrossExpDesigns.qmd

### Type I errors across experimental designs {#extra}
We examine the generalizability of the preceding results across alternative experimental designs involving two or three factors. We assess five of the six ratio-scale distributions considered earlier, excluding the Cauchy distribution and replacing it with a 5-level ordinal scale with flexible thresholds.

**Main effects**. @fig-designs-main illustrates Type I error rates for the main effect of $X_2$ while varying the magnitude of the main effect of $X_1$, focusing on $n = 20$. RNK and INT consistently maintain error rates close to $5\%$, with the exception of the Poisson distribution, for which error rates become low in the two within-subjects designs (particularly for INT). Similarly, PAR generally maintains error rates near $5\%$, but they drop to notably low levels for certain combinations of designs ($2 \times 2 \times 2$ and $3 \times 3 \times 3$ within-subjects) and distributions (log-normal and exponential).

In contrast, ART inflates error rates across all non-normal distributions, with the extent of inflation varying by design. For example, ART performs particularly poorly for the ordinal scale when applied to the $2 \times 2 \times 2$ and $3 \times 3 \times 3$ designs. In these designs, we also observe inflated error rates for log-normal data, even when $a_1 = 0$.

::: {#fig-designs-main}
```{r, echo=FALSE, fig.height=3.3, fig.width = 9, warning=FALSE}
prefix <- "Type_I_designs"
alpha = 0.05

distributions = c("norm", "lnorm", "exp", "poisson", "binom", "likert")
dnames = c("Normal", "Log-normal", "Exponential", "Poisson", "Binomial", "Ordinal (5 levels)")

df <- read_data(prefix, alpha, effectType = 1, distributions)
df <- reshape_by_design(df, dnames)

plotly_error_by_design(df, xlab = "magnitude of main effect", var = "rateX2", xvar = "effectX1", max = 101)
```
Type I error rates ($\alpha = .05$) for the **main effect of $X_2$** as a function of the magnitude $a_1$ of the main effect of $X_1$ ($n = 20$)
:::

**Interaction effects**. @fig-designs-interaction-1 displays Type I error rates for the interaction effect $X_1 \times X_2$ while varying the effect of $X_1$ ($n = 20$). We observe consistent trends in line with our previous findings. For additional insights, we direct readers to our raw experimental data, which demonstrate that these trends persist across other interaction terms, namely $X_1 \times X_3$, $X_2 \times X_3$, and $X_1 \times X_2 \times X_3$, within the two 3-factor designs.

::: {#fig-designs-interaction-1}
```{r, echo=FALSE, fig.height=3.3, fig.width = 9, warning=FALSE}
plotly_error_by_design(df, xlab = "magnitude of main effect", var = "rateX1X2", xvar = "effectX1", max = 101)
```
Type I error rates ($\alpha = .05$) for the **interaction** $X_1 \times X_2$ as a function of the magnitude $a_1$ of the main effect of $X_1$ ($n = 20$) 
:::